State and Parameter Estimation for a Spatio-Temporal Neural Fields

### 2010-04-14 ltx/Journal of Neuroengineering/StateAndParameterEstimationNeuralFields_JoNE.tex
updated parameter list

 The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \lambda\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\lambda$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the dynamic model
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{k +t}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^-=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace N-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\end{algorithmic}
\end{algorithm}
\section{Comparison of Estimated Parameters and True Parameters}

\section{Discussion}
\begin{itemize}

@@

\section*{References}
\bibliographystyle{unsrt}
\bibliography{BrainIDE}

>>>>

 The accuracy of the approximation is linked to the time step used in the temporal discretization of the system, where the state transition appears more linear with a finer resolution. To explain the operation of the UKF we begin by deriving the traditional Kalman filter with the state-space model
\begin{eqnarray}
	\mathbf{x}_t = q(\mathbf{x}_t)\boldsymbol{\theta} - \lambda\mathbf{x}_t + k\left(u_t\left(\mathbf{r}\right)\right) + \mathbf{e}_t \\
	\mathbf{y}_t = \mathbf{C}\mathbf{x}_t + \boldsymbol{\varepsilon}_t,
\end{eqnarray}
where $\boldsymbol{\theta}$ and $\lambda$ are the systems parameters. In forwards iteration since the observation equation is linear in parameters we only need to use Unsented transformation in prediction step. The update step equations remains the same as the standard Kalman filtering formulations. In backward smoothing path, estimation from the forward filtering is used to form the new set of sigma points in a similar way to the filterin step, however the predicted cross-covariance of the states need to be calculated. A summary of the algorithm is given in Algorithm 
\begin{algorithm}
\caption{The Unscented Kalman Filter}\label{CPTEMDAlgorithm}
\begin{algorithmic}[1]
\State Initialise
\begin{equation*}
\hat{\mathbf x}_0, \mathbf P_0
\end{equation*}
\State Forward iteration
For $t \in \left\lbrace 0,\cdots, N\right\rbrace $,
\begin{equation*}
\mathcal X_t^f=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^{f}-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
Propogate the sigma points through the dynamic model
\begin{equation}
\mathcal X_{i,t+1|t}^{f*}=Q(\mathcal X_t^f,u_t)
\end{equation}
\begin{equation*}
 \hat{\mathbf x}_{k +t}^{f-}=\sum_{i=0}^{2L} W_i^{(m)}\mathcal X_{i,t+1|t}^{f*}
\end{equation*}
\begin{equation*}
 \mathbf P_{t +1}^-=\sum_{i=0}^{2L} W_i^{(c)}(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})(\mathcal X_{i,t+1}^{f-}-\hat{\mathbf x}_{t +1}^{f-})^\top+\boldsymbol \Sigma_e
\end{equation*}
\begin{equation*}
\mathcal K_{t+1}=\mathbf P_{t +1}^{f-}\mathbf C ^\top(\mathbf C \mathbf P_{t +1}^{f-}\mathbf C ^\top+\boldsymbol \Sigma_v)^{-1}
\end{equation*}
\begin{equation*}
\hat{\mathbf x}_{t+1}^{f}=\hat{\mathbf x}_{t+1}^{f-}+\mathcal K_{t+1}(\mathbf y_{t+1}-\mathbf C\hat{\mathbf x}_{t +1}^{f-})
\end{equation*}
\begin{equation*}
 \mathbf P_{t+1}^f=(\mathbf I - \mathcal K_{t+1}\mathbf C)\mathbf P_{t +1}^{f-}
\end{equation*}
\State Backward iteration
\State initialisation
\begin{equation}
 \mathbf P_T^s= \mathbf P_T^f
\end{equation}
\begin{equation}
\hat{\mathbf x}^s_T= \hat{\mathbf x}^f_T
\end{equation}
For $t \in \left\lbrace N-1, \cdots, 0 \right\rbrace $
\begin{equation*}
\mathcal X_t^s=[\hat{\mathbf x}_t^f \quad \hat{\mathbf x}_t^f+\gamma\sqrt{\mathbf P_t^f} \quad \hat{\mathbf x}_t^f-\gamma\sqrt{\mathbf P_t^f}]
\end{equation*}
\end{algorithmic}
\end{algorithm}
\section{Results}

\subsection{Frequency Analysis}

\subsection{State and Parameter Estimates}

\section{Discussion}

\begin{itemize}

@@

\section{Parameters}

\begin{tabular} {c|c} 
	\hline\hline
	Symbol & Description \\ \hline
	$v(\mathbf{r},t)$ & mean membrane potential \\
	$g(\mathbf{r},t)$ & average action potential rate \\
	$\mathbf{r}$ & spatial location \\
	$t$ & time (s) \\
	$h(t)$ & post-synaptic response kernel \\
	$\eta(t)$ & Heaviside function \\
	$\alpha$ & inverse synaptic time constant \\
	$w(\mathbf{r},\mathbf{r}')$ & spatial connectivity kernel \\
	$f(\mathbf{r},t)$ & firing function rate \\
	$u(\mathbf{r},t)$ & external input \\
	$\Omega$ & spatial domain \\
	$f_{max}$ & maximal firing rate \\
	$\beta$ & slope of sigmoidal activation function \\
	$v_0$ & firing threshold \\
	$\delta(t)$ & Dirac-delta function \\
	$L$ & temporal differential operator \\
	$e(\mathbf{r},t)$ & field disturbance, covariance $\gamma$\\
	$m(\mathbf{r},\mathbf{r}')$ & sensor kernel, variance $\sigma_m^2$ \\
	$\epsilon(\mathbf{r}_n,t)$ & observation noise, covariance $\Sigma_\epsilon$ \\
	$y(\mathbf{r}_n,t)$ & obseration \\
	$n$ & sensor index $n=1,...,N$ \\
	$T$ & time step \\
	$\lambda$ & time constant parameter \\
	$\boldsymbol{\phi(r)}$ & vector of Gaussian basis functions \\
	$\mathbf{x}_t$ & state vector at time $t$ \\
	$\boldsymbol{\psi}$ & vector of connectivity kernel basis functions \\
	$\boldsymbol{\theta}$ & vector of connectivity kernel parameters \\
	$\boldsymbol{\Gamma}$ & inner product of field basis functions \\
	$q(.)$ & state transition function \\
	$k(.)$ & maps into to discrete state of input function $u(\mathbf{r},t)$ \\
	$\mathbf{e}_t$ & state disturbance, covariance $\Sigma_e$ \\
	$\mathbf{C}$ & observation matrix \\
	$V(\nu)$ & Spectrum of the dynamic field \\ 
	$\nu$ & Spatial frequency \\
	$\nu_c$ & Spatial cut-off frequency \\
\end{tabular}

\begin{tabular}{c|c}
	\hline\hline
	Symbol & Description \\ \hline
	$\Delta_s$ & Distance between adjacent sensors \\
	$\rho$ & Over-sampling parameter \\	
	$\Delta_b$ & Distance between field basis functions \\	
	$\sigma_{nu}^2$ & Variance of FT of Gaussian basis function \\
	$\sigma^2$ & Spatial variance of Gaussian in spatial domain \\
	$\hat{\mathbf{x}}$ & State estimate \\
	$\boldsymbol{\chi}$ & Matrix of sigma vectors \\
	$\hat{\mathbf{x}}_t^{f-}$ & Forward \emph{a priori} estimate \\
	$\hat{\mathbf{x}}_t^f$ & Forward \emph{a posteriori} estimate \\
	$\hat{\mathbf{x}}_t^{b-}$ & Backward \emph{a priori} estimate \\
	$\hat{\mathbf{x}}_t^{b}$ & Backward \emph{a posteriori} estimate \\
	$\mathbf{P}^f_t$ & Forward \emph{a posteriori} covariance matrix \\
	$\mathbf{P}^{f-}_t$ & Forward \emph{a priori} covariance matrix \\	
\end{tabular}
\section*{References}
\bibliographystyle{unsrt}
\bibliography{BrainIDE}
